What\,Agreement\,Hath\,Mars\,With\,Earth? \\ An Exploration of Urbit's IPC Interfaces

### 2024-04-25 mss.tex
Final cut.

Formally, Nock (and thus Urbit) is functional-as-in-language and free of side effects.  In practice, the \texttt{hint} opcode 11 in Nock is used to pass some noun to the interpreter.  The interpreter can then choose a higher order action to take based on this value, whether I/O, memory management, or other side effects.  This allows for the \texttt{king} to pass hints to the \texttt{serf} which produces effects on Earth (send a network packet, get a keyboard input, etc.).  This bridging of the Earth/Mars gap allows for vanes to interact outside of Mars.

@@

\section[{\texttt{\%lick}:  The Low-Level {\sc ipc} Interface}]{\texttt{\%lick}:  The Low-Level IPC Interface}

@@

Gall needs to wrap \lstinline[style=inlinecode]{%soak} and \lstinline[style=inlinecode]{%spit} to route properly (see e.g. \lstinline[style=inlinecode]{++ap-generic-take}).  This permits multiple agents to share sockets with the same name while still organizing an agent's sockets by a unique path.

@@

This exploration into Urbit's {\sc ipc} interfaces, in particular \texttt{\%khan}, \texttt{conn.c}, and \texttt{\%lick}, has sought to illuminate the architectural innovations that distinguish Urbit from traditional computing paradigms.  These mechanisms permit Vere to balance Arvo's demand for pristine determinism against the host Unix OS's many competing software interfaces and tools.  While Urbit itself obviously relies deeply on these, following this pattern in other systems could lead to a redefinition of how digital ecosystems interact. \tombstone{}

>>>>

Formally, Nock (and thus Urbit) is functional-as-in-language and free of side effects.  In practice, the \lstinline[style=inlinecode]{%hint} opcode 11 in Nock is used to pass some noun to the interpreter.  The interpreter can then choose a higher order action to take based on this value, whether I/O, memory management, or other side effects.  This allows for the \texttt{king} to pass hints to the \texttt{serf} which in turn produces effects on Earth (sends a network packet, retrieves keyboard input, etc.).  This bridging of the Earth/Mars gap allows for vanes to interact outside of Mars.

@@

\section[{\texttt{\%lick}:  The Low-Level \textsc{\textbf{ipc}} Interface}]{\texttt{\%lick}:  The Low-Level IPC Interface}

@@

Gall needs to wrap \lstinline[style=inlinecode]{%soak} and \lstinline[style=inlinecode]{%spit} to route properly (see e.\,g.\ \lstinline[style=inlinecode]{++ap-generic-take}).  This permits multiple agents to share sockets with the same name while still organizing an agent's sockets by a unique path.

@@

\sloppy
This exposition of Urbit's {\sc ipc} interfaces, in particular \texttt{\%khan}, \texttt{conn.c}, and \texttt{\%lick}, has sought to cast light upon architectural innovations that distinguish Urbit from traditional computing paradigms.  These mechanisms permit Vere to balance Arvo's demand for pristine determinism against the host Unix OS's many competing software interfaces and tools.  While Urbit itself obviously relies deeply on these, following this pattern in other systems could lead to a redefinition of how digital ecosystems interact. \tombstone{}
